---
title: "lab 8"
author: "Riyesh Nath"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
install.packages("standardize")
```

```{r, echo=FALSE}
library(ggplot2)
library(tidyverse)
library(haven)
library(pROC)
library(standardize)

setwd("~/Documents/Masters/Econometrics/data")# your directory structure will be different
load("ACS_2021_couples.RData")

#summary(acs2021_couples)
```

```{r, echo=FALSE}

acs2021_couples$RACE <- fct_recode(as.factor(acs2021_couples$RACE),
                                      "White" = "1",
                                      "Black" = "2",
                                      "American Indian or Alaska Native" = "3",
                                      "Chinese" = "4",
                                      "Japanese" = "5",
                                      "Other Asian or Pacific Islander" = "6",
                                      "Other race" = "7",
                                      "two races" = "8",
                                      "three races" = "9")
                                      
acs2021_couples$h_race <- fct_recode(as.factor(acs2021_couples$h_race),
                                   "White" = "1",
                                   "Black" = "2",
                                   "American Indian or Alaska Native" = "3",
                                   "Chinese" = "4",
                                   "Japanese" = "5",
                                   "Other Asian or Pacific Islander" = "6",
                                   "Other race" = "7",
                                   "two races" = "8",
                                   "three races" = "9")

acs2021_couples$HISPAN <- fct_recode(as.factor(acs2021_couples$HISPAN),
                                     "Not Hispanic" = "0",
                                     "Mexican" = "1",
                                     "Puerto Rican" = "2",
                                     "Cuban" = "3",
                                     "Other" = "4")
acs2021_couples$h_hispan <- fct_recode(as.factor(acs2021_couples$h_hispan),
                                     "Not Hispanic" = "0",
                                     "Mexican" = "1",
                                     "Puerto Rican" = "2",
                                     "Cuban" = "3",
                                     "Other" = "4")

```

#### In this homework, we attempt to see what parameters we can use to find if a given Male household head is 10 years of older than his Female partner.

```{r}
trad_data <- acs2021_couples %>% filter( (SEX == "Female") & (h_sex == "Male") )
trad_data$he_more_than_10yrs_than_her <- as.numeric(trad_data$age_diff <= -5)
```

#### We first check to see if interracion couple versus same race couple grouped by race has an affect on the proportion of women who are marrier to men older than 10 years. Looking at the plot below, we do see that race and couple type (interracial or same race) does have an affect.


```{r}
trad_data$same_race <- as.numeric(trad_data$RACE == trad_data$h_race)
race_interracial <- trad_data %>%
  mutate(couple_type = ifelse(same_race == 1, "Same Race", "Interracial")) %>%
  group_by(RACE, couple_type) %>%
  summarize(
    prop_10plus_gap = mean(he_more_than_10yrs_than_her),
    n = n()
  ) %>%
  filter(n >= 100) 

ggplot(race_interracial, aes(x = reorder(RACE, prop_10plus_gap), y = prop_10plus_gap, fill = couple_type)) +
  geom_col(position = "dodge") +
  geom_text(aes(label = paste0(round(prop_10plus_gap * 100, 1), "%")), 
            position = position_dodge(width = 0.9),
            hjust = -0.1, size = 3) +
  coord_flip() +
  labs(title = "Proportion of Couples Where Man is 10+ Years Older",
       subtitle = "By Woman's Race and Interracial Status",
       x = "Woman's Race",
       y = "Proportion",
       fill = "Couple Type") +
  scale_y_continuous(labels = scales::percent, limits = c(0, 0.23)) +
  scale_fill_manual(values = c("Interracial" = "#E74C3C", "Same Race" = "#3498DB")) +
  theme_minimal() +
  theme(legend.position = "top")

```
#### Looking at the proportion of partnership where men are older women based on education level of men and woman, we see similar pattern where we see a lower education level of both men and women affecting the age difference. Though the proportion isn't too significant.

```{r}

trad_data <- trad_data %>% filter(RACE == "Other Asian or Pacific Islander" | RACE == "Chinese" | RACE == "Black")
unique(trad_data$RACE)
trad_data$RACE <- fct_drop(trad_data$RACE)
```

```{r}


education_status_of_women_who_married_older_men_df <- as.data.frame.matrix(xtabs(data = trad_data, formula = ~  EDUC + he_more_than_10yrs_than_her))
education_status_of_women_who_married_older_men_df <- rownames_to_column(education_status_of_women_who_married_older_men_df, var="EDUC")

education_status_of_women_who_married_older_men_df$total <- education_status_of_women_who_married_older_men_df$`0` + education_status_of_women_who_married_older_men_df$`1`
education_status_of_women_who_married_older_men_df$prop <- education_status_of_women_who_married_older_men_df$`1` / education_status_of_women_who_married_older_men_df$total
education_status_of_women_who_married_older_men_df <- education_status_of_women_who_married_older_men_df %>%
  filter(!is.na(prop))

ggplot(data= education_status_of_women_who_married_older_men_df, aes(x = prop, y = EDUC)) +
  geom_col() + 
  labs(title = "Proportion of Couples Where Man is 10+ Years Older",
       subtitle = "By Woman's Education",
       y = "Woman's Education",
       x = "Proportion")



education_status_of_men_who_married_younger_women_df <- as.data.frame.matrix(xtabs(data = trad_data, formula = ~  h_educ + he_more_than_10yrs_than_her))
education_status_of_men_who_married_younger_women_df <- rownames_to_column(education_status_of_men_who_married_younger_women_df, var="h_educ")

education_status_of_men_who_married_younger_women_df$total <- education_status_of_men_who_married_younger_women_df$`0` + education_status_of_men_who_married_younger_women_df$`1`
education_status_of_men_who_married_younger_women_df$prop <- education_status_of_men_who_married_younger_women_df$`1` / education_status_of_men_who_married_younger_women_df$total
education_status_of_men_who_married_younger_women_df <- education_status_of_men_who_married_younger_women_df %>%
  filter(!is.na(prop))

ggplot(data= education_status_of_men_who_married_younger_women_df, aes(x = prop, y = h_educ)) +
  geom_col() + 
  labs(title = "Proportion of Couples Where Man is 10+ Years Older",
       subtitle = "By Men's Education",
       y = "Man's Education",
       x = "Proportion")


```

#### We also look at the proportion of those born in the US state against those not born in the US state. We see that a higher proportion of those not born in the US state have 10 years age gap between man and woman. However, here the proportion both both are still very low. 

```{r}

proportion_based_on_born_in_usa <- as.data.frame.matrix(xtabs(data= trad_data, formula = ~ born_in_USstate + he_more_than_10yrs_than_her))
proportion_based_on_born_in_usa <- rownames_to_column(proportion_based_on_born_in_usa, var="born_int_USstate")
proportion_based_on_born_in_usa$total <- proportion_based_on_born_in_usa$`0` + proportion_based_on_born_in_usa$`1`
proportion_based_on_born_in_usa$prop <- proportion_based_on_born_in_usa$`1` / proportion_based_on_born_in_usa$total

ggplot(data = proportion_based_on_born_in_usa, aes(y = born_int_USstate, x = prop)) +
  geom_col() +
  labs(
    title = "Proportion of Couples Where Man is 10+ Years Older",
       subtitle = "By born in US",
       y = "BORN IN",
       x = "Proportion"
  )



```
#### We also see that difference states may have an affect the proportions overall are very minor.

```{r}
by_state <- as.data.frame.matrix(xtabs(data= trad_data, formula = ~ STATEFIP + he_more_than_10yrs_than_her))

by_state <- rownames_to_column(by_state, var="State")
by_state$total <- by_state$`0` + by_state$`1`
by_state$prop <- by_state$`1` / by_state$total

by_state$State <- with(
  by_state,
  reorder(State, prop)
)

ggplot(data= by_state, aes(x= prop, y= State)) + 
  geom_col()
```

```{r}
using_data <- trad_data
using_data$State_factor <- as.factor(using_data$STATEFIP)
#levels(using_data$STATEFIP)

d_y_varb <- data.frame(model.matrix(~ using_data$he_more_than_10yrs_than_her)) 

d_State <- data.frame(model.matrix(~ using_data$State_factor))

print(d_y_varb)
sum( colSums(d_State) == 0)
```

#### The overall trend that we see with low proportion we see in all the graphs above may be attributed to overall low percent of Men who have married women 10 years or younger. 

```{r}
ggplot(trad_data, aes(y = he_more_than_10yrs_than_her)) +
  geom_bar(aes(x = after_stat(prop))) +
  labs(
    title = "Proportion of Couples Where Man is 10+ Years Older",
    x = "Proportion",
    y = "He is More Than 10 Years Older"
  )
```

#### Below are steps to create design matrix for each feature and then put them together for 

```{r}
data_used_for_our_design_matrix <- trad_data

data_used_for_our_design_matrix$EDUC <- fct_drop(data_used_for_our_design_matrix$EDUC)

design_matrix_age_diff <- data.frame(model.matrix( ~ data_used_for_our_design_matrix$he_more_than_10yrs_than_her))
design_matrix_same_race <- data.frame(model.matrix(~ data_used_for_our_design_matrix$same_race))
design_matrix_RACE <- data.frame(model.matrix(~ data_used_for_our_design_matrix$RACE))
design_matrix_h_race <- data.frame(model.matrix(~ data_used_for_our_design_matrix$h_race))
design_matrix_EDUC <- data.frame(model.matrix(~ data_used_for_our_design_matrix$EDUC))
design_matrix_STATEFIP <- data.frame(model.matrix(~ data_used_for_our_design_matrix$STATEFIP))
print(design_matrix_STATEFIP)

```

```{r}
sum(colSums(design_matrix_same_race) == 0 )
sum(colSums(design_matrix_RACE) == 0)
sum(colSums(design_matrix_h_race) == 0)
sum(colSums(design_matrix_EDUC) == 0)
sum(colSums(design_matrix_STATEFIP) == 0)
sum(colSums(design_matrix_age_diff) == 0)

```

```{r}

design_matrix_for_logit <- data.frame(
  design_matrix_age_diff[,2],
  design_matrix_same_race[,2], 
  design_matrix_RACE[,2:length(design_matrix_RACE)],
  design_matrix_h_race[,2:length(design_matrix_h_race)],
  design_matrix_EDUC[,2:length(design_matrix_EDUC)],
  design_matrix_STATEFIP[,2:length(design_matrix_STATEFIP)]
)

names(design_matrix_for_logit)[1] <- "he_more_than_10yrs_than_her"


names(design_matrix_for_logit) <- sub("data_used_for_our_design_matrix.","",names(design_matrix_for_logit))

names(design_matrix_for_logit)
```

```{r}
require("standardize")
set.seed(654321)

NN <- length(design_matrix_for_logit$he_more_than_10yrs_than_her)
print(NN)

restrict_1 <- (runif(NN) < 0.05) # use 10% as training data, ordinarily this would be much bigger but start small
summary(restrict_1) # you should understand wtf this variable is for
data_train <- subset(design_matrix_for_logit, restrict_1)
data_test <- subset(design_matrix_for_logit, !restrict_1)

# again check this below, should be zero
sum( colSums(data_train) == 0)


```
```{r}
#print(design_matrix_for_logit)
formula_sobj <- reformulate( names(design_matrix_for_logit[2:length(design_matrix_for_logit)]), response = "he_more_than_10yrs_than_her")

sobj <- standardize(formula_sobj, data_train, family = binomial) # standardized object

s_dat_test <- predict(sobj, data_test)

print(sobj$data)
```

```{r}
# OLS linear probability model
model_lpm1 <- lm(sobj$formula, data = sobj$data)
summary(model_lpm1)
pred_vals_lpm <- predict(model_lpm1, s_dat_test)
pred_model_lpm1 <- (pred_vals_lpm > mean(pred_vals_lpm))
table(pred = pred_model_lpm1, true = data_test$he_more_than_10yrs_than_her )

# logit 
model_logit1 <- glm(sobj$formula, family = binomial, data = sobj$data)
summary(model_logit1)
pred_vals <- predict(model_logit1, s_dat_test, type = "response")
pred_model_logit1 <- (pred_vals > mean(pred_vals) )
table(pred = pred_model_logit1, true = data_test$he_more_than_10yrs_than_her )
```

```{r}
logistic_modal <- glm(data = trad_data, formula = he_more_than_10yrs_than_her ~ same_race + RACE + h_race + EDUC + STATEFIP , family = binomial)

predict_val <- predict(logistic_modal, trad_data, type = "response")

roc_obj <- roc(trad_data$he_more_than_10yrs_than_her, predict_val)

best_thr <- coords(roc_obj, "best", ret = "threshold")
print(best_thr)

summary(logistic_modal)
auc(roc_obj)
pred <- as.integer(predict_val >= 0.110481)
table(pred, trad_data$he_more_than_10yrs_than_her)


```

```{r}
require('randomForest')
set.seed(54321)
model_randFor <- randomForest(as.factor(he_more_than_10yrs_than_her) ~ ., data = sobj$data, importance=TRUE, proximity=TRUE)
print(model_randFor)
round(importance(model_randFor),2)
varImpPlot(model_randFor)
# look at confusion matrix for this too
pred_model1 <- predict(model_randFor,  s_dat_test)
table(pred = pred_model1, true = data_test$he_more_than_10yrs_than_her)

```
