---
title: "Homework#5"
author: "Riyesh Nath"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

**Group Members:**
**Riyesh Nath, Jason S., Bamba Cisse and Romulo Jiminez**

### Summary
**In our experiment, we attempt to create a model that allows us to predict potential income_midpoint given certain parameters from household census data. For our model, we will use linear regression model.**

```{r, echo=FALSE}
library(tidyverse)
library(ggplot2)
library(viridis)
library(modelsummary)

setwd("~/Documents/GIT/Econometrics/Homework#3")
load("data/d_HHP2020_24.Rdata")
```

### Data Exploration
**For the first step, we want to examine the parameters that might be useful for our linear regression. Lets examine by doing some data exploration. We will focus on a subcategory, using NorthEast Region**

```{r, echo=FALSE}
## Function definitions - Guys, we can just reuse so no need for excess code. 

draw_box_plot <- function(data, column){
  ggplot(data, 
         aes(x = .data[[column]], y = income_midpoint, color= .data[[column]])
         ) +
    geom_boxplot() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
}

draw_anova <- function(data, column){
  formula <- as.formula(paste("income_midpoint~", column))
  anov <- aov(formula, data = data)
  summary(anov)
}

draw_scatter_plot <- function(data, column){
  ggplot(data, 
         aes(x= .data[[column]], y=income_midpoint, color= .data[[column]])
         ) +
    geom_point() +
    geom_smooth(method = "lm", se = TRUE)
}

```

### Use ANOVA for categorical data to see which one has a greater effect on income_midpoint

```{r, echo=FALSE}
d_HHP2020_24 <- d_HHP2020_24 %>% filter(Region=="Northeast")

list_of_variables <- c("Education", "Mar_Stat", "Race", "Hispanic", "private_health_ins", 
                       "work_kind", "public_health_ins", "Census_division", "Gender", "workloss")

## I iterate through a loop of the list of variables so we don't have to recoden this. And inside it I do whatever I want to.
## right now I have a box plot and anova to see which has significant value
## I removed NA.

for(item in list_of_variables){
  filter_data = d_HHP2020_24 %>% filter(!is.na(.data[[item]]))
  print(draw_box_plot(filter_data, item))
  print(draw_anova(filter_data, item))
}
```

### Use geom_point to see which continuous data has a greate effect on income_midpoint

```{r, echo=FALSE}
list_of_scatter_var <- c("Age", "Number_people_HH", "Number_kids_HH", "ANXIOUS", "WORRY", "INTEREST", "DOWN")
for(item in list_of_scatter_var){
  filter_data = d_HHP2020_24 %>% filter(!is.na(.data[[item]]))
  print(draw_scatter_plot(filter_data, item))
}
```


```{r, echo=FALSE}

filtered_data <- d_HHP2020_24 %>% filter(
    !is.na(Education) & 
    !is.na(workloss) &
    !is.na(Gender) &
    !is.na(Mar_Stat) &
    !is.na(Race) &
    !is.na(Number_people_HH) &
    !is.na(Number_kids_HH) &
    !is.na(private_health_ins) &
    !is.na(public_health_ins) &
    Age >= 25 & 
    Age <= 55 
  )

model_1 <- lm(income_midpoint ~ Education + Mar_Stat + Race + private_health_ins + 
                public_health_ins + Gender + workloss +
                Number_people_HH + Number_kids_HH, 
              data = filtered_data)

summary(model_1)
```
