---
title: "Homework#5"
author: "Riyesh Nath"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

**Group Members:**
**Riyesh Nath, Jason Seda, and Bamba Cisse**

### Summary
**In our experiment, we attempt to create a model that allows us to predict potential income_midpoint given certain parameters from household census data. For our model, we will use linear regression model.**

```{r, echo=FALSE}
library(tidyverse)
library(ggplot2)
library(viridis)
library(modelsummary)

setwd("~/Documents/GIT/Econometrics/Homework#3")
load("data/d_HHP2020_24.Rdata")
```

### Data Exploration
**For the first step, we want to examine the parameters that might be useful for our linear regression. Lets examine by doing some data exploration. We will focus on a subcategory, using NorthEast Region**

```{r, echo=FALSE}
## Function definitions - Guys, we can just reuse so no need for excess code. 

draw_box_plot <- function(data, column){
  ggplot(data, 
         aes(x = .data[[column]], y = income_midpoint, color= .data[[column]])
         ) +
    geom_boxplot() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
}

draw_anova <- function(data, column){
  formula <- as.formula(paste("income_midpoint~", column))
  anov <- aov(formula, data = data)
  summary(anov)
}

draw_scatter_plot <- function(data, column){
  ggplot(data, 
         aes(x= .data[[column]], y=income_midpoint, color= .data[[column]])
         ) +
    geom_point() +
    geom_smooth(method = "lm", se = TRUE)
}

```

### Use ANOVA for categorical data to see which one has a greater effect on income_midpoint

```{r, echo=FALSE}
d_HHP2020_24 <- d_HHP2020_24 %>% filter(Region=="Northeast")

list_of_variables <- c("Education", "Mar_Stat", "Race", "private_health_ins", "public_health_ins",
                       "Gender", "workloss")

## I iterate through a loop of the list of variables so we don't have to recoden this. And inside it I do whatever I want to.
## right now I have a box plot and anova to see which has significant value
## I removed NA.

for(item in list_of_variables){
  filter_data = d_HHP2020_24 %>% filter(!is.na(.data[[item]]))
  print(draw_box_plot(filter_data, item))
  print(draw_anova(filter_data, item))
}
```

**Income Midpoint comparison vs. education:**

There is a clear and consistent positive correlation between education and income. For 
individuals with schooling up to the associate degree level, income tends to rise gradually, 
showing a modest but steady upward trend. However, after completing a bachelor’s degree, 
there is a substantial and more pronounced increase in income levels. This sharp jump 
highlights a significant differentiation between those with a college education and those with 
only some postsecondary or associate-level attainment, emphasizing the strong economic 
advantage associated with higher educational attainment.

**Income vs. marital status:**

There is no substantial separation in income levels across most categories, except for a clear 
and notable increase among married individuals. This jump is largely attributed to the presence 
of multiple earners within the same household, which naturally raises the combined household 
income. Apart from this distinction, income differences among other groups remain relatively 
moderate and do not exhibit significant divergence.

**Income vs. Race:**

While, in principle, income should not differ by race or ethnicity, the data indicate a clear 
tendency for higher income levels among individuals identifying as Asian or White in the 
Northeast. This pattern may be influenced by cultural norms, educational attainment, and 
occupational distribution prevalent within these groups. Indirectly, these factors contribute to 
reinforcing existing socioeconomic stereotypes, even if the underlying causes are more 
structural than individual.

**Income vs. private Health insurance obtained:**

Those with private health insurance consistently show higher overall median income levels.

**Income vs. public health insurance:**

A similar pattern appears when comparing public health insurance to income levels, suggesting 
that lower-income individuals are generally less likely to hold any form of health insurance.

**Income vs. Gender:**

It shows a general income differentiation of male vs all “others”, yet, trans gender shows a much
lower differentiation between those of the trans category.

**Income vs. Workloss:**

As expected, those with a recent loss in work have a lower overall expected income midpoint.

### Use geom_point to see which continuous data has a greate effect on income_midpoint

```{r, echo=FALSE}
list_of_scatter_var <- c( "ANXIOUS")
for(item in list_of_scatter_var){
  filter_data = d_HHP2020_24 %>% filter(!is.na(.data[[item]]))
  print(draw_scatter_plot(filter_data, item))
}
```

**Income vs Anxiety:**

When plotting general anxiety levels against income midpoints, we see a gradual decline in 
income as anxiety increases. This could suggest that financial instability and job insecurity 
contribute to higher anxiety, or conversely, that chronic anxiety limits access to higher-paying 
roles due to challenges in high-pressure environments. It's also likely the relationship is cyclical, 
anxiety both results from and reinforces lower income. This underscores the need to view 
mental health as a key factor in economic opportunity.

```{r, echo=FALSE}

filtered_data <- d_HHP2020_24 %>% filter(
    !is.na(Education) & 
    !is.na(Mar_Stat) &
    !is.na(Race) &
    !is.na(private_health_ins) &
    !is.na(public_health_ins) &
    !is.na(Gender) &
    !is.na(workloss) &
    !is.na(ANXIOUS) &
    Age >= 25 & 
    Age <= 55 
  )

model_1 <- lm(income_midpoint ~ Education + Mar_Stat + Race + private_health_ins + 
                public_health_ins + Gender + workloss + ANXIOUS,
              data = filtered_data)

summary(model_1)
```
